EDIpack2: interoperable Lanczos-based solver for generic quantum impurity problems

### 2025-05-13 05_examples.tex
First draft of the edipack2triqs example subsection

\section{Examples}\label{SecExamples}
In this section we illustrate the functioning of the \NAME
library and their interfaces through specific examples, including
discussion of relevant code snippets.
Specifically, we show how the quantum impurity solver in \NAME can be used
to address problems of different nature within the framework of DMFT.

\subsection{Bethe lattice DMFT}\label{SecExamplesBetheDMFT}
The description of the Mott or metal to insulator transition (MIT)
within the Bethe lattice, i.e. Cayley tree with connection
$z\to\infty$ is conventionally considered the test bed of any DMFT
application. Here  we discuss a guided implementation based on the
Fortran API of \NAME of the DMFT solution of the Bethe lattice at
$T=0$. 
We consider a Fermi-Hubbard model:
$$
H = -t \sum_{\langle ij\rangle\s}c^+_{i\s}c_{j\s} + U \sum_i n_{i\up}n_{i\dw}
$$
where $c^+_{i\s}$ ($c_{i\s}$) is the creation (destruction) operator for an electron at site $i$
and with spin $\s$, $n_{i\s}=c^+_{i\s}c_{i\s}$ is the occupation
operator and the first sum is over the nearest neighbor sites $\langle
ij\rangle$. 
The system is defined on a Bethe lattice with density of states
$\rho(\e)=\frac{1}{2D}\sqrt{D^2-\e^2}$, where $D=2t$ is the
half-bandwidth~\cite{Georges1996}.
In DMFT this model is mapped
onto a quantum impurity problem with an effective electronic bath
which needs to be determined self-consistently.
\begin{lstlisting}[style=fstyle,numbers=none,basicstyle={\scriptsize\ttfamily}]
program ed_hm_bethe
   USE EDIPACK2
   USE SCIFOR
   implicit none
   !> Number of discretization points for the Bethe DOS
   integer,parameter                           :: Le=5000
   !> Bethe half-bandwidth = energy unit
   real(8),parameter                           :: D=1d0
   !>Bath:
   real(8),allocatable                         :: Bath(:)
   !> local dynamical functions, rank-5 [Nspin,Nspin,Norb,Norb,L]
   complex(8),allocatable,dimension(:,:,:,:,:) :: Weiss,Smats,Sreal,Gmats,Greal,Weiss_
   ... (omissis)
   !> READ THE input using EDIpack procedure:
   call ed_read_input(trim(finput))
   !
   !> Setup the Bethe DOS $\rho$ (linear dispersion)
   Ene = linspace(-D,D,Le,mesh=de)
   DOS = dens_bethe(Ene,D)
\end{lstlisting}

The bath is described
by the (unknown) function $\GG^{-1}_0$, i.e. the Weiss field. Within \NAME
exact diagonalization approach, where the bath is approximated with a
finite number of levels, the function $\GG^{-1}_0$ is used to
determine bath parameters $\vec{x}=\{V,h\}$ using the optimization method discussed in
\secu{sSecFit}.
The starting point of any calculation is determined by a suitable
guess of the Weiss field, or equivalently of the bath parameters.
In \NAME this operations is accomplished by the function {\tt
  ed\_init\_solver}.

\begin{lstlisting}[style=fstyle,numbers=none,basicstyle={\scriptsize\ttfamily}]
Nb=ed_get_bath_dimension()
allocate(bath(Nb))
call ed_init_solver(bath)
\end{lstlisting}

The iterative algorithm proceeds as follows:
\begin{enumerate}
\item Call the exact diagonalization {\bf impurity solver} {\tt
    ed\_solve} whose only input is the set of parameters $\vec{x}$. All other
  options are controlled by input file specifications ({\tt
    ed\_read\_input}).
\item Retrieve at least the self-energy functions $\Sigma_{\a\b\s\s'}(i\omega)$ on the
  Matsubara axis using {\tt ed\_get\_sigma}.
\item Evaluate the local interacting Green's function
  $G_\mathrm{loc}(i\omega) = \int_{-D}^D \frac{\rho(\e)}{\zeta -\e}d\e$ with
  $\zeta=i\omega+\mu-h^0-\Sigma(i\omega)$.
  \item Update the Weiss field through the {\bf self-consistency}
    relation: $\GG^{-1}_0(i\omega) = G_\mathrm{loc}(i\omega) + \Sigma(i\omega)$. 
  \item Optimize the bath parameters $\vec{x}$ against the updated
    Weiss field using, for instance, {\tt ed\_chi2\_fitgf}. Restart at 1.
\end{enumerate}
In agreement with Reverse Communication Strategy setup, the step 3-4
are left to the user. In addition step 5 is kept separated to
guarantee enough freedom in the bath optimization, though we remark
this is a crucial step which can potentially compromise the whole
calculation.
An example of implementation is given in the following listings.

\begin{lstlisting}[style=fstyle,numbers=none,basicstyle={\scriptsize\ttfamily}]
iloop=0;converged=.false.
   do while(.not.converged.AND.iloop<nloop)
     iloop=iloop+1     
     !> Solve the effective impurity problem
     call ed_solve(bath)     
     !> Retrieve impurity Self-energy on Matsubara axis
     call ed_get_sigma(Smats,'m')
     !
     !> Build a local Green's function:
     wfreq = pi/beta*(2*arange(1,Lmats)-1)   !automatic Fortran allocation
     do i=1,Lmats ! One can do better than this of course 
        zeta= xi*wfreq(i)+xmu - Smats(1,1,1,1,i)
        Gmats(1,1,1,1,i) = sum(DOS(:)/( zeta-Ene(:) ))*de  
     enddo
     !> Update Weiss field: Self-consistency
     Weiss(1,1,1,1,:) = one/(one/Gmats(1,1,1,1,:) + Smats(1,1,1,1,:))
     !> Close the self-consistency fitting the new bath:
     call ed_chi2_fitgf(Weiss,bath,ispin=1)     
     !>Check convergence
     converged =( sum(abs(Weiss(1,1,1,1,:)-Weiss_(1,1,1,1,:)))/sum(abs(Weiss(1,1,1,1,:))) )<dmft_error
     Weiss_=Weiss     
   enddo
\end{lstlisting}

\paragraph{Results}
In the following we present some results obtained for the model
introduced above. Specifically we focus on the testbed case of the interaction
driven Mott transition, descriving the progressive
transformation of a partially-filled band metallic state into a correlated insulator.

To illustrate this point, we report in panel (A) the evolution of
the spectral function $-{\rm Im}G_\mathrm{loc}(\omega)/\pi$ as a function of
the interaction strenght $U$. Despite the {\it spiky} nature of the spectrum, due to the
finite size (i.e. number of poles) of the discretized effective bath,
it is possible to clearly distinguish the characteristic features of
the Mott transition. At low energy, we observe the renormalization of the 
quasi-particle peak at the Fermi level ($\omega=0$). Concomitantly the
system shows the formation of incoherent  high-energy features which
will then develop into Hubbard bands in the Mott insulating state for $U>U_\mathrm{c}$, with $U_\mathrm{c}\simeq 2.8D$.

@@

The formation of a spectral gap separating the Hubbard bands in the
Mott insulator is supported by the divergence, at the Fermi level, of the imaginary part of
the self-energy function. Causality imposes that the real-part of
this quantity becomes large near the singularity making it impossible
to satisfy the quasi-particles poles equations
$\omega+\mu-h^0-\e-\Re{\Sigma}(\omega)=0$, i.e. preventing coherent
quasi-particles excitations near the Fermi level. 
In the panels (B) and (C) we discuss this point  showing the evolution
of the self-energy functions $\Sigma$.
Specifically, in panel (B) we report the self-energy ${\rm Im}\Sigma(i\omega)$
on the Matsubara axis in the low energy regime. Increasing $U$
we observe the progressive growth of this function until a
diverging behavior is acheived crossing the critical interaction
strenght $U>U_\mathrm{c}$.
We recall that this behavior can be observed on the Matsubara axis
because of the particle-hole symmetry of the problem which pins the
$\Im{\Sigma}$ singularity at $\omega=0$.
In panel (C), we report the corresponding behavior of the
$\Re{\Sigma}(\omega)$ in small energy regime around the Fermi
level. Augmenting the interaction strength $U$ leads to a sudden
increase of this self-energy component until a discontinuous behavior
is reached, associated to the divergence of the corresponding
imaginary part.

A quantitative description of the transition is obtained from the
quasi-particle renormalization constant $Z$, which correlates with the degree of
localization of the electrons being equal to 1 for a non-interacting
metal and 0 for the Mott insulator. This quantity is obtained from the
linear term in the Fermi liquid expansion of the self-energy as
$Z=(1-\tfrac{\partial\Re\Sigma}{\partial\omega}_{|_{\omega\rightarrow
    0}})^{-1}$.
Using the relation:
$$
   \frac{\Im\Sigma(i\omega_n)}{\omega_n}_{|_{\omega_n\rightarrow 0}}=
   \frac{1}{\pi}\int_{\mathbb R}d\epsilon \frac{\Re\Sigma(\epsilon)}{\epsilon^2}=
   \frac{\partial\Re\Sigma}{\partial\omega}_{|_{\omega\rightarrow 0}}.
$$
we can extract $Z$ also from the linear behavior of ${\rm Im}\Sigma(i\omega)$ for 
$\omega\to0$ in the metallic regime.

The linear fit highlighted in the panels (B)-(C) and the corresponding values of $Z$ are reported in the legend.
In both cases we observe that increasing $U$ the slope of the linear behavior of the
self-energy at low-energy increases upon increasing $U$, either on the
Matsubara or real-axis. At the transition point the slope becomes
infinite according to the divergence of the imaginary part of
the self-energy  ${\rm Im}\Sigma(\omega) \rightarrow -\infty$,
corresponding to $Z\to 0$.

\subsection{Attractive Hubbard model (Python API)}\label{SecExamplesAHM}
The second application has the twofold aim to demonstrate the
superconductivity support of the \NAME library as well as to
illustrate the use of the Python API through a working example.
We consider the simple, yet non-trivial, case of the attractive
Hubbard model on a two-dimensional square lattice:

$$
H = \sum_{\ka\s} \e(\ka)c^+_{\ka\s}c_{\ka\s} -U\sum_i n_{i\up}n_{i\dw}
$$
where $U>0$, $c^+_{\ka\s}=\tfrac{1}{\sqrt{N}}\sum_i e^{-i\ka\cdot R_i} c^+_{i\s}$
and $c^+_{i\s}$ is the creation operator for an electron at site $i$
and with spin $\s$, $n_{i\s}=c^+_{i\s}c_{i\s}$ is the occupation
operator and $\e(\ka)=-2t[\cos{(k_x a)}+ \cos{(k_y a)}]$ is the
energy dispersion relation. In the following we the lattice distance
$a=1$  and the energy unit as $4t=D=1$.

The DMFT workflow for this case is very similar to the previous one
provided works in the Nambu basis defined by the spinor
$\psi_i=[c_{i\up}, c^+_{i\dw}]^t$.
In this basis any correlation functions, e.g. the Green's function, takes the matrix form:
\begin{equation}
  {\mathbf G} =
  \begin{pmatrix}
    \hat{G}_{\uparrow\uparrow} & \hat{F}_{\uparrow\downarrow}\\
    \hat{\bar{F}}_{\downarrow\uparrow}  &    \hat{\bar{G}}_{\downarrow\downarrow} \\
   \end{pmatrix}
\end{equation}
where the symbol $\hat{A}$ indicates the possible multi-orbital
character which reduced to a scalar in the present case of a single
orbital. The symbol $\hat{\bar{A}}$ indicates that the components in
the second row are connect by particle-hole and time-reversal
symmetries to the first row one. The specific relation depends on the
symmetry of the order parameter (here $s$-wave) and the nature of the
complex argument of the functions, i.e. Matsubara or real-axis:
\begin{equation}
\begin{array}{cc}
  \hat{\bar{G}}(i\omega) = -\hat{G}^*(i\omega)\;; &  \hat{\bar{F}}(i\omega) = \hat{F}(i\omega)\\
  \hat{\bar{G}}(\omega)  = -\hat{G}^*(-\omega) \;; & \hat{\bar{F}}(\omega) = \hat{F}^*(i\omega)\\
\end{array}
\end{equation}  
The code implementation closely follows the previous structure, with
some notable exceptions related to the Nambu structure for which,
using the symmetries above, we only need to deal with the first row
independent components. In the first part of the code we initialize
the lattice structure and the solver:
\begin{lstlisting}[style=mypython,numbers=none,basicstyle={\scriptsize\ttfamily}]
import numpy as np
from edipy2 import global_env as ed
import mpi4py
from mpi4py import MPI
import os,sys
from aux_funx import * #include function to build $G_\mathrm{loc}$ and Weiss fields

comm = MPI.COMM_WORLD
rank = comm.Get_rank()
master = (rank==0)

#Functions
def generate_kgrid(Nk):
    b1=2*np.pi*np.array([1.0,0.0])
    b2=2*np.pi*np.array([0.0,1.0])
    n1, n2 = np.meshgrid(np.arange(Nk), np.arange(Nk))
    n1=n1/Nk;n2=n2/Nk
    gridout = np.stack([n1.ravel(), n2.ravel()], axis=-1)
    return np.dot(gridout,[b1,b2])

def h_square2d(k,t):
  return -2*t*(np.cos(k[...,0,np.newaxis,np.newaxis])+np.cos(k[...,1,np.newaxis,np.newaxis]))*np.eye(ed.Norb)

ed.read_input("inputAHM.conf")

kgrid = generate_kgrid(Nk)

#Generate hk and hloc
Hk   = h_square2d(kgrid,t_hop)
HkNambu   = np.array([h_square2d(kgrid,t_hop),-np.conj(h_square2d(-kgrid,t_hop))])
Hloc = np.sum(Hk,axis=0)/Nk**2
Hloc = Hloc.astype(complex)
ed.set_hloc(Hloc)

#SETUP SOLVER
Nb=ed.get_bath_dimension()
bath = ed.init_solver()
bath_prev = np.copy(bath)
\end{lstlisting}

>>>>

\section{Examples}\label{SecExamples}
In this section we illustrate the functioning of the \NAME
library and their interfaces as solver for DMFT through specific examples, including
discussion of relevant code snippets.

\subsection{Bethe lattice DMFT}\label{SecExamplesBetheDMFT}
The test bed for any DMFT application is conventionally considered 
the description of the Mott or metal-to-insulator transition (MIT) 
in a Bethe lattice, i.e., a Cayley tree with coordination number 
$z\to\infty$~\cite{Georges1996}.
Here, we present a guided implementation of the DMFT 
solution for the Bethe lattice at $T=0$ using \NAME as impurity
solver.
The model under consideration is the Fermi-Hubbard Hamiltonian:
$$
H = -t \sum_{\langle ij\rangle,\s} c^\dagger_{i\s} c_{j\s} + 
    U \sum_i n_{i\up}n_{i\dw}
    $$

where $c^\dagger_{i\s}$ ($c_{i\s}$) are the creation (annihilation) 
operators for an electron at site $i$ with spin $\s$, and 
$n_{i\s} = c^\dagger_{i\s} c_{i\s}$ is the corresponding occupation 
operator. The first sum runs over nearest-neighbor pairs 
$\langle ij \rangle$.

The system is defined on a Bethe lattice with density of states
$\rho(\e)=\frac{1}{2D}\sqrt{D^2-\e^2}$,
where $D=2t$ is the half-bandwidth.
In DMFT this lattice model is mapped onto a quantum impurity problem with 
an effective electronic bath that must be determined
self-consistently.

Below, we present the key components of a basic DMFT implementation 
using \NAME for the Bethe lattice, employing either the Fortran or 
C++ APIs. Fully operational versions of the two codes can be found 
in the {\tt examples} directory of the \NAME source code. Both 
samples share a similar initial structure, including memory 
allocation, Bethe lattice creation, and solver initialization.

\setbox0=\hbox{%
  \begin{minipage}{0.465\linewidth}
    \center{\bf Fortran}
\begin{lstlisting}[style=fstyle,frame=none,numbers=none,basicstyle={\scriptsize\ttfamily}]
program ed_hm_bethe
   USE EDIPACK2
   USE SCIFOR
   implicit none
   integer                :: Le=1000
   real(8)                :: wmixing=0.5d0
   real(8)                :: D=1d0
   integer                :: Nb
   real(8),allocatable    :: Bath(:)
   complex(8),allocatable :: Hloc(:,:,:,:)
   real(8),allocatable    :: Ebands(:),Dbands(:)
   complex(8),allocatable :: Smats(:,:,:,:,:)
   complex(8),allocatable :: Delta(:,:,:,:,:)
   ...  
   !Read variables
   call ed_read_input('inputED.conf')
   !Solver-specific arays
   allocate(Smats(Nspin,Nspin,Norb,Norb,Lmats))
   allocate(Delta(Nspin,Nspin,Norb,Norb,Lmats))
   allocate(Hloc(Nspin,Nspin,Norb,Norb));Hloc=0d0
   ...
   !Initialize Bethe arrays
   allocate(Ebands(Le),Dbands(Le))
   Ebands = linspace(-D,D,Le,mesh=de)
   Dbands = dens_bethe(Ebands,Wband)*de
   !Set local H
   call ed_set_hloc(Hloc)
\end{lstlisting}
\end{minipage}
}
\savestack{\listingA}{\box0}

\setbox0=\hbox{%
  \begin{minipage}{0.505\linewidth}
    \center{\bf C++}
\begin{lstlisting}[style=cstyle,frame=none,numbers=none,basicstyle={\scriptsize\ttfamily}]
#include <edipack2_cbinding.h>
using namespace std;
...
//Main variables
int Le = 1000;
int iloop = 0;
double wmixing = 0.5;
double D = 1.0;
//Read  variables    
char input[] = "inputED.conf"; 
read_input(input);      
//Dimensions
int64_t d[4] = {Nspin,Nspin,Norb,Norb};
int total_size = d[0] * d[1] * d[2] * d[3];
int total_size_n5 = total_size * Lmats;    
// Solver-specific arrays
vector<complex<double>> Hloc(total_size);
vector<complex<double>> Smats(total_size_n5);
vector<complex<double>> Delta(total_size_n5);
...
// Initialize Bethe arrays
vector<double> Ebands, Dbands;
Ebands = linspace(-D,D,Le);
Dbands = dens_bethe(Ebands,Wband,de);
//set local H
ed_set_Hloc_single_N4(Hloc.data(), d);
\end{lstlisting}
\end{minipage}
}
\savestack{\listingB}{\box0}

\begin{tabular}{c|c}\label{list2}
  \stackinset{l}{}{t}{}{}{\listingA} & \stackinset{l}{}{t}{}{}{\listingB} 
\end{tabular}

The bath is described by the (unknown) function $\GG^{-1}_0$, known as 
the Weiss field ({\tt Weiss}). In the ED approach implemented in \NAME, 
the bath is approximated using a finite number of discrete energy levels. 
The function $\GG^{-1}_0$ is then used to determine the bath parameters 
$\vec{x} = \{V, h\}$ through the optimization method outlined in 
\secu{sSecFit}.

The starting point for any calculation is a reasonable initial guess 
for the Weiss field, or equivalently, the bath parameters. In \NAME, 
this is accomplished using the function {\tt
  ed\_init\_solver} (Fortran API) or {\tt init\_solver\_site} (C++ API).

\setbox0=\hbox{%
  \begin{minipage}{0.465\linewidth}
    \center
    \begin{lstlisting}[style=fstyle,frame=none,numbers=none,basicstyle={\scriptsize\ttfamily}]
   !Initialize solver
   Nb=ed_get_bath_dimension()
   allocate(bath(Nb))
   call ed_init_solver(bath)
\end{lstlisting}
\end{minipage}
}
\savestack{\listingA}{\box0}

\setbox0=\hbox{%
  \begin{minipage}{0.505\linewidth}
    \center
\begin{lstlisting}[style=cstyle,frame=none,numbers=none,basicstyle={\scriptsize\ttfamily}]
// Initialize solver
int Nb;
vector<double> Bath(Nb);
Nb = get_bath_dimension();
int64_t bath_dim[1] = {Nb};
init_solver_site(Bath.data(), bath_dim);
\end{lstlisting}
\end{minipage}
}
\savestack{\listingB}{\box0}

\begin{tabular}{c|c}\label{list2}
  \stackinset{l}{}{t}{}{}{\listingA} & \stackinset{l}{}{t}{}{}{\listingB} 
\end{tabular}

The iterative algorithm to solve the DMFT problem is based on
a reverse communication strategy and proceeds as follows:
\begin{itemize}

\item[{\tiny $\blacksquare$}] Call the exact diagonalization {\bf impurity solver}
  whose only input is the set of parameters $\vec{x}$ contained in a
  rank-1 array. All  \NAME options are controlled through input file
  specifications.

\item[{\tiny $\blacksquare$}] Retrieve the self-energy functions $\Sigma_{\a\b\s\s'}(i\omega)$ on the
  Matsubara axis using dedicated function provided by \NAME API.

\item[{\tiny $\blacktriangleright$}] Evaluate the local interacting Green's function
  $G_\mathrm{loc}(i\omega) = \int_{-D}^D \frac{\rho(\e)}{\zeta -\e}d\e$ with
  $\zeta=i\omega+\mu-h^0-\Sigma(i\omega)$.

\item[{\tiny $\blacktriangleright$}] Update the Weiss field via the {\bf self-consistency}
  relation: $\GG^{-1}_0(i\omega) = G_\mathrm{loc}(i\omega) +
    \Sigma(i\omega)$. For the Bethe lattice, this simplifies to
    $\Delta = \tfrac{D}{4}G_\mathrm{loc}$.

  \item[{\tiny $\triangleright$}] Optimize the bath parameters $\vec{x}$ against the updated
    Weiss field,  potentially using the conjugate gradient procedures 
supplied by \NAME. Then, restart at step 1.
\end{itemize}

The first two steps, marked by the ${\tiny \blacksquare}$ symbol, 
are handled directly by \NAME routines. In the subsequent steps, 
indicated by the ${\tiny \blacktriangleright}$ symbol, control returns 
to the user, who must implement the algebraic updates required to 
close the self-consistency loop and optimize the bath.
Given the 
critical importance of this optimization, \NAME provides access to 
well-tested functions, ensuring stability and reproducibility of 
results. Alternative optimization methods can also be employed as 
needed. 
An example of implementation is provided in the following listings.

\setbox0=\hbox{%
  \begin{minipage}{0.45\linewidth}
    \center
    \begin{lstlisting}[style=fstyle,frame=none,numbers=none,basicstyle={\scriptsize\ttfamily}]
!DMFT loop
do while(.not.converged.AND.iloop<nloop)
    iloop=iloop+1     
    ! Call ED Solver
    call ed_solve(bath)     
    ! Retrieve Self-energy on Matsubara axis
    call ed_get_sigma(Smats,'m')
    ! Build local Green's function
    wfreq = pi/beta*(2*arange(1,Lmats)-1)
    do i=1,Lmats
       zeta= xi*wfreq(i)+xmu - Smats(1,1,1,1,i)
       Gmats(1,1,1,1,i)=sum(DOS/(zeta-Ene))*de  
    enddo
    ! Self-consistency
    Delta = D/4d0*Gmats
    ! Fitting -> new bath
    call ed_chi2_fitgf(Weiss,bath,ispin=1)     
    ! Check convergence
    converged=check_convergence(Delta,&
          dmft_error,Nsuccess,Nloop)
enddo
\end{lstlisting}
\end{minipage}
}
\savestack{\listingA}{\box0}

\setbox0=\hbox{%
  \begin{minipage}{0.52\linewidth}
    \center
\begin{lstlisting}[style=cstyle,frame=none,numbers=none,basicstyle={\scriptsize\ttfamily}]
  // DMFT loop
  while (iloop < Nloop && !converged) {
    //  Call ED Solver
    solve_site(Bath.data(),bath_dim,1,1);      
    // Retrieve  Self-energy on Matsubara axis
    get_sigma_site_n5(Smats.data(),//
        0,0,wm.data(),Lmats,0);
    //  Build a local Green's function:
    for (int i=0;i<Lmats;++i) {
      zeta= wm[i] + xmu - Smats[i];
      Gmats[i] = complex<double>(0.0,0.0);
      for (int j=0; j< Le; j++) {
        Gmats[i]+=Dbands[j]/(zeta-Ebands[j]);
      }
    }
    // Self-consistency
    for (int i = 0; i < Lmats; ++i) {
      Delta[i] = 0.25 * D * Gmats[i];
    }
    // Fit -> new bath
    chi2_fitgf_single_normal_n5(Delta.data(),//
        delta_dim,Bath.data(),bath_dim,1,0,1);
    // Check convergence
    converged = check_convergence(Delta,//
        dmft_error, Nsuccess, Nloop, comm);  
  }
\end{lstlisting}
\end{minipage}
}
\savestack{\listingB}{\box0}

\begin{tabular}{c|c}\label{list2}
  \stackinset{l}{}{t}{}{}{\listingA} & \stackinset{l}{}{t}{}{}{\listingB} 
\end{tabular}

\paragraph{Results}
In the following, we present results for the interaction-driven Mott 
transition in the Bethe lattice, which captures the gradual 
transformation of a partially-filled metallic state into a correlated 
insulator.
To illustrate this, we report in panel (A) the evolution of the spectral 
function $-{\rm Im}G_\mathrm{loc}(\omega)/\pi$ as a function of the 
interaction strength $U$. Despite the inherently {\it spiky} nature 
of the spectrum, resulting from the finite number of poles in the 
discretized effective bath, the characteristic features of the Mott 
transition are clearly visible. At low energies, a renormalized 
quasi-particle peak develops at the Fermi level ($\omega = 0$). 
Simultaneously, the system exhibits the formation of incoherent 
high-energy features, which eventually evolve into well-defined 
Hubbard bands in the Mott insulating phase for $U > U_\mathrm{c}$, 
with $U_\mathrm{c} \simeq 2.8D$.

@@

The formation of a spectral gap separating the Hubbard bands in the 
Mott insulating phase is associated with the divergence of the 
imaginary part of the self-energy at the Fermi level. This divergence 
reflects the complete localization of the electrons, effectively 
suppressing coherent quasi-particle excitations. Specifically, 
causality dictates that the real part of the self-energy must also 
grow significantly near the singularity, making it impossible to 
satisfy the quasi-particle pole equation
$$
\omega+\mu-h^0-\e-\Re{\Sigma}(\omega)=0,
$$
which governs the formation of coherent excitations near the Fermi 
level.

Panels (B) and (C) illustrate this phenomenon by showing the evolution 
of the self-energy $\Sigma$. In panel (B), we present the Matsubara 
self-energy ${\rm Im}\Sigma(i\omega)$ in the low-energy regime. As the 
interaction strength $U$ increases, this function progressively 
grows, eventually diverging as the critical interaction threshold 
$U > U_\mathrm{c}$ is crossed.
This divergence along the Matsubara axis is directly linked to the
particle-hole symmetry of the Bethe lattice, which pins the  ${\rm
  Im}\Sigma$ singularity at $\omega = 0$.

Panel (C) complements this picture by displaying the real part 
$\Re{\Sigma}(\omega)$ in the same low-energy regime. Here, increasing 
$U$ leads to a rapid rise of this component, culminating in a 
discontinuous behavior as the critical point is approached. This 
discontinuity directly reflects the divergence in the imaginary part
on the real-axis,  confirming the transition to the Mott insulating state.

A quantitative measure of this transition is provided by the 
quasi-particle renormalization factor $Z$, which can be used to capture the degree 
of electron delocalization. This parameter ranges from 1 for a 
non-interacting metal to 0 for a fully localized Mott insulator. It is 
defined through the low-energy expansion of the self-energy as
$$
Z=(1-\tfrac{\partial\Re\Sigma}{\partial\omega}_{|_{\omega\rightarrow
    0}})^{-1}.
$$
which can also be estimated from the linear behavior of ${\rm Im}\Sigma(i\omega)$ for 
$\omega\to0$ in the metallic regime using the relation:
$$
   \frac{\Im\Sigma(i\omega_n)}{\omega_n}_{|_{\omega_n\rightarrow 0}}=
   \frac{1}{\pi}\int_{\mathbb R}d\epsilon \frac{\Re\Sigma(\epsilon)}{\epsilon^2}=
   \frac{\partial\Re\Sigma}{\partial\omega}_{|_{\omega\rightarrow 0}}.
$$

The linear fits highlighted in panels (B) and (C), along with the 
corresponding $Z$ values provided in the legends, clearly indicate that 
the slope of the self-energy at low energy increases with $U$ on both 
the Matsubara and real axes. At the transition point, this slope 
diverges, reflecting the onset of complete electron localization as 
$Z \to 0$, consistent with the divergence 
${\rm Im}\Sigma(\omega) \rightarrow -\infty$.

\subsection{Attractive Hubbard model}\label{SecExamplesAHM}
The second example we present concerns the DMFT description of the 
attractive Hubbard model on a two-dimensional square lattice. This 
example has a twofold aim: (i) to demonstrate the support of \NAME 
for ($s$-wave) superconductivity, and (ii) to illustrate the use of 
the library's Python API through a concrete working example. The 
model Hamiltonian is given by:
$$
H = \sum_{\ka,\s} \epsilon(\ka) c^\dagger_{\ka\s} c_{\ka\s} 
    - U \sum_i n_{i\up} n_{i\dw},
$$
where $U > 0$, $c^\dagger_{\ka\s} = \tfrac{1}{\sqrt{N}} 
\sum_i e^{-i\ka \cdot R_i} c^\dagger_{i\s}$, and $c^\dagger_{i\s}$ 
is the creation operator for an electron at site $i$ with spin $\s$. 
The occupation operator is $n_{i\s} = c^\dagger_{i\s} c_{i\s}$, and 
the energy dispersion relation is $\e(\ka)=-2t[\cos{(k_x a)}+
\cos{(k_y a)}]$, where we set the lattice spacing 
$a=1$  and the choose the energy unit such as $4t=D=1$ for convenience.

The DMFT workflow for this case is largely similar to the previous 
example, but it now operates in the Nambu basis defined by the spinor
$$
\psi_i=\begin{bmatrix} c_{i\up} \\ c^\dagger_{i\dw} \end{bmatrix}. 
$$
In this basis, the Green's function takes the matrix form:
\begin{equation}
  {\mathbf G} =
  \begin{pmatrix}
    \hat{G}_{\uparrow\uparrow} & \hat{F}_{\uparrow\downarrow}\\
    \hat{\bar{F}}_{\downarrow\uparrow}  &    \hat{\bar{G}}_{\downarrow\downarrow} \\
  \end{pmatrix}
\end{equation}
where the symbol $\hat{A}$ indicates the potential multi-orbital 
nature of the system, which reduces to a scalar in the present 
single-orbital case. The components in the second row, denoted 
by $\hat{\bar{A}}$, are connected to the first row by particle-hole 
and time-reversal symmetries. The specific relations depend on the 
symmetry of the order parameter (here, $s$-wave) and whether the 
functions are defined on the Matsubara or real-frequency axis:
\begin{equation}
\begin{array}{cc}
  \hat{\bar{G}}(i\omega) = -\hat{G}^*(i\omega)\;; &  \hat{\bar{F}}(i\omega) = \hat{F}(i\omega)\\
  \hat{\bar{G}}(\omega)  = -\hat{G}^*(-\omega) \;; & \hat{\bar{F}}(\omega) = \hat{F}^*(i\omega)\\
\end{array}
\end{equation}

The code implementation closely follows the structure of the previous 
example, with some notable adjustments related to the Nambu basis. 
Specifically, the above symmetries allow one to compute only the 
independent components from the first row, significantly reducing 
the computational effort. The initial part of the code handles the 
lattice structure and solver initialization, as described below.

\begin{lstlisting}[style=mypython,numbers=none,basicstyle={\scriptsize\ttfamily}]
import numpy as np
from edipy2 import global_env as ed
import mpi4py
from mpi4py import MPI
import os,sys
from aux_funx import * #include function to build $G_\mathrm{loc}$ 
comm = MPI.COMM_WORLD
rank = comm.Get_rank()
master = (rank==0)
#Functions
def generate_kgrid(Nk):
    b1=2*np.pi*np.array([1.0,0.0])
    b2=2*np.pi*np.array([0.0,1.0])
    n1, n2 = np.meshgrid(np.arange(Nk), np.arange(Nk))
    n1=n1/Nk;n2=n2/Nk
    gridout = np.stack([n1.ravel(), n2.ravel()], axis=-1)
    return np.dot(gridout,[b1,b2])
def h_square2d(k,t):
  return -2*t*(np.cos(k[...,0,np.newaxis,np.newaxis])+np.cos(k[...,1,np.newaxis,np.newaxis]))*np.eye(ed.Norb)

#Read input
ed.read_input("inputAHM.conf")

#Generate Hk and  set Hloc
kgrid = generate_kgrid(Nk)
Hk   = h_square2d(kgrid,t_hop)
HkNambu   = np.array([h_square2d(kgrid,t_hop),-np.conj(h_square2d(-kgrid,t_hop))])
Hloc = np.sum(Hk,axis=0)/Nk**2
ed.set_hloc(Hloc.astype(complex))

#Setup ED Solver
Nb=ed.get_bath_dimension()
bath = ed.init_solver()
\end{lstlisting}

The iterative scheme for the solution of DMFT closely follows the
sequence already discussed in \secu{SecExamplesBetheDMFT}:   
\begin{itemize}
\item Call the exact diagonalization {\bf impurity solver} {\tt
    ed.solve} providing the set of bath parameters $\vec{x}=\{V,h\}$  as input.

\item Use the dedicated
  input/output \NAME procedures to retrieve the self-energy functions  
  $\hat{\Sigma}(i\omega)$ and $\hat{S}(i\omega)$ on the 
  Matsubara axis.

\item
  Evaluate the interacting local Green's functions $\hat{G}_\mathrm{loc}$ and
  $\hat{F}_\mathrm{loc}$:
  \begin{equation}
  {\mathbf G}_\mathrm{loc}(i\omega) =
  \int_\RRR d\e \rho(\e)
  \begin{pmatrix}
    (i\omega +\mu)\hat{\11} - \hat{h}^0 - \hat{\Sigma}(i\omega) -\e & -\hat{S}(i\omega) \\
    -\hat{S}(i\omega) & (i\omega +\mu)\hat{\11} + \hat{h}^0 +
    \hat{\Sigma}^*(i\omega) +\e\\
  \end{pmatrix}^{-1}
\end{equation}

\item Update the Weiss fields components, respectively 
  $\GG_0^{-1}$ and $\FF_0^{-1}$, through the {\bf self-con\-sis\-ten\-cy}
  relation: $\mathbfcal{G}^{-1}_0(i\omega) = {\mathbf G}_\mathrm{loc}(i\omega) +
  {\mathbf \Sigma}(i\omega)$ in Nambu space.

\item Optimize the bath parameters $\vec{x}$ to best describe the updated
    Weiss fields, potentially using \NAME provided conjugate gradient  fit
    procedures.
  \end{itemize}
%
The corresponding implementation in Python reads:
\begin{lstlisting}[style=mypython,numbers=none,basicstyle={\scriptsize\ttfamily}]
#DMFT CYCLE
converged=False;iloop=0
while (not converged and iloop<ed.Nloop ):
    iloop=iloop+1
    #Solve impurity problem
    ed.solve(bath)    
    #Self-consistency with aux_funx procedures:
    Smats = np.array([ed.get_sigma(axis="m",typ="n"),ed.get_sigma(axis="m",typ="a")])    
    Gmats = get_gloc(wm*1j       ,ed.xmu,HkNambu,Smats,axis="m") 
    Weiss = dmft_weiss_field(Gmats,Smats)                            
    #Fit
    bath = ed.chi2_fitgf(Weiss[0],Weiss[1],bath)
    #Error check
    err,converged=ed.check_convergence(Weiss,ed.dmft_error)
ed.finalize_solver()
\end{lstlisting}
